---
title: "HW4"
output: pdf_document
---
Question 1 Preparation Codes
```{r}

library(ISLR)
library(boot)
library(margins)
library(glmnet)
library(dplyr)
library(tidyr)

set.seed(317)
getwd()
setwd("/Users/nishengwenxin/Desktop")
train = read.csv("gss_train.csv")%>%
  na.omit

test = read.csv("gss_test.csv")%>%
  na.omit
```
# Question 1

Find the optimal degree d.

```{r}
seed = 317
cv.MSE<- rep(NA, 10)
for (i in 1:10) {
    glm.fit <- glm(egalit_scale ~ poly(income06, i), data = train)
    cv.MSE[i] <- cv.glm(train, glm.fit, K=10)$delta[1]}

plot( x = 1:10, y = cv.MSE, xlab = "power of income", ylab = "10-fold CV error", 
      main = 'Choosing the optimal degree d (1.a)',
      type = "b", pch = 19, lwd = 2, bty = "n", 
      ylim = c( min(cv.MSE) - sd(cv.MSE), max(cv.MSE) + sd(cv.MSE) ) )

points( x = which.min(cv.MSE), y = min(cv.MSE),
        col = "red", pch = "X", cex = 1.5 )

```

From diagram 1.a, we can see that the the relationship between the responsible variable $Y_{egalit_scale}$ and the predictive variable $X_{income06}$ in the gss-training data set is best modeled by the polynomial regression of degree 2. Equivalently, we get the optimal model with the expression: $$Y= \beta_0 + \beta_1 X +\beta_2 X^2 + \epsilon $$

It's worth to notice that $d=2$ is an acceptable value because generally, it is unusual to use d greater than 3 or 4 because for large values of d, the polynomial curve can become overly flexible and can take on some very strange shapes.

The diagram below (diagram 1.b) shows the resulting polynomial fit to the data and the AME (diagram 1.c)
```{r}
range.income06 <- range(train$income06)
income06.grid <- seq(from = range.income06[1], to = range.income06[2])

fit <- lm(egalit_scale ~ stats::poly(income06, 2), data = train)
preds <- predict(fit, newdata = list(income06 = income06.grid))
plot(egalit_scale ~ income06, 
     main = 'The polynomial fit with the optimal degree (1.b)',
     ylab = 'Predicted egalitarian scale ',
     data = train, col = "blue")
lines(income06.grid, preds, col = "red", lwd = 2)

cplot(fit,"income06",what = "effect",
      main= "AME for income06 (1.c)")
```



# Question 2
Choose the optimal number of cuts. 
```{r}
cv.MSE <- rep(NA, 10)
# for each cut perform 10-fold cross-validation
for (i in 2:10) {
  train$income06.cut <- cut(train$income06, i)
  lm.fit <-  glm(egalit_scale ~ income06.cut, data = train)
  cv.MSE[i] <-  cv.glm(train, lm.fit, K = 10)$delta[1]
}
  plot(2:10, cv.MSE[-1], xlab = "Number of cuts", ylab = "10-fold CV Error",
       main = 'Choosing the optimal number of cuts (2.a)',
     type = "b", pch = 19, lwd = 2, bty ="n")

  points( x = which.min(cv.MSE), y = min( cv.MSE, na.rm = TRUE), col = "red", pch = "X", cex = 1.5 )

```
 
 Diagram 2.a shows that the test error is minimized when the number of cuts equals 4. Equivalently, we breaks X (income06) into 4 distinctive bins and pick a different constant for each of these 4 bins. Since there exists dummy varaibles, we in fact get 4+1 = 5 distinctive regions. The mathematical expression for this model is $$Y=\beta_0 + \beta_1 C_1(X) + \beta_2 C_2(X) + ...+\beta_4 C_4(X) + \epsilon$$

The diagram below (diagram 2.b) shows the resulting fit with the 4 number of cuts.

```{r}

fit  <- glm(egalit_scale~ cut(income06, 4), data = train)
preds <- predict(fit, data.frame(income06 = income06.grid))

plot(egalit_scale ~ income06, data = train, col = "blue",
     main = 'The step function with the optimal number of cuts (2.b)',
     ylab = 'Predicted egalitarian scale ')
lines(income06.grid, preds, col = "red", lwd = 2)

```

# Question 3 

Select the optimal number of degrees of freedom


```{r}
library(splines)
set.seed(317)
cv.MSE <- rep(NA, 10)
for (i in 3:10) {
    fit <- glm(egalit_scale ~ ns(income06, df = i), data = train)
    cv.MSE[i] <- cv.glm(train, fit, K = 10)$delta[1]
}

plot(3:10, cv.MSE[-c(1, 2)], xlab = "Degrees of freedom", ylab = "10-fold CV error", 
     main = 'Choosing the optimal degree of freedom (3.a)',type = "l")
d.min <- which.min(cv.MSE)
points(which.min(cv.MSE), cv.MSE[which.min(cv.MSE)], col = "red", cex = 2, pch = 20)
```



Diagram 3.a shows that the test error is minimized when the degree of freedom equals 6. This spline model is a continuous degree-6 polynomial with continuity in
derivatives up to degree 5 at each knot. In addition, since it's a natural spline, the function is linear at the tail, in order to combat the danger of high variance at the outer range of the predictors.

The diagram below (diagram 3.b) shows the resulting fit with the 6 df.

```{r}

fit <- glm(egalit_scale ~ ns(income06, df = 6), data = train)
preds <- predict(fit, data.frame(income06 = income06.grid))

plot(egalit_scale ~ income06, data = train, col = "blue",
     ylab = 'Predicted egalitarian scale ',
     main = 'Natural spline with the optimal degree of freedom (3.b)')
lines(income06.grid, preds, col = "red", lwd = 2)

```

# Question 4

## 4.a Linear Regression
```{r}
library(caret)

train = read.csv("gss_train.csv")
train.num = train %>% select(authoritarianism,childs,con_govt,egalit_scale,
                              income06,science_quiz,sibs,social_connect,
                              tolerance,tvhours,wordsum)
train.num1 = train%>%
    mutate_if(is.numeric, c) %>%
    mutate_if(is.numeric, scale) %>%
    mutate_if(is.matrix, as.numeric)


preprocessParams <- preProcess(select_if(train, is.numeric), 
                               method=c("center", "scale"))

train_trans <- predict(preprocessParams, select_if(train, is.numeric))
test = read.csv("gss_test.csv")
test.num = test %>% select(authoritarianism,childs,con_govt,egalit_scale,
                              income06,science_quiz,sibs,social_connect,
                              tolerance,tvhours,wordsum)


lm.fit=lm(egalit_scale~.,data=train.num)

pred<-predict(lm.fit,test.num)
print(lm.fit)
cat('MSE Value:', mean((pred-test$egalit_scale)^2))
```

```{r}

train.mat <- model.matrix(egalit_scale ~ ., data = train.num)
test.mat <- model.matrix(egalit_scale ~ ., data = test.num)
for (i in seq(0, 1, .1))
{
  seed = 828
  cv.out <- cv.glmnet (train.mat,train$egalit_scale,alpha = i)
  bestlam <- cv.out$lambda.min
  model <- glmnet(train.mat,train$egalit_scale,alpha=1,
                  lambda=bestlam)
  pred <- predict(model,s = bestlam ,newx = test.mat)
  MSE <- mean((pred-test$egalit_scale)^2)
  cat('lambda:',bestlam,'alpha:',i,'MSE:',MSE,"\n")
}
```
Therefore，the optimal model occurs where alpha is 0.4

```{r}
cv.out <- cv.glmnet (train.mat,train$egalit_scale,alpha = 0.4)
bestlam <- cv.out$lambda.min
net.model <- glmnet(train.mat,train$egalit_scale,alpha=1,
                  lambda=bestlam)
pred <- predict(net.model,s = bestlam ,newx = test.mat)
MSE <- mean((pred-test$egalit_scale)^2)
net.model
cat('MSE for Elastic Net:',MSE)

```

## 4.c Principal component regression
```{r}
library(pls)
pcr_model <- pcr(egalit_scale~., data = train.num, scale = TRUE, validation = "CV")
validationplot(pcr_model,val.type = 'MSEP',
               main = 'PCR Model and its MSE (4.c)',
               ylab = 'MSE by cross-validation')
summary(pcr_model)
```
Diagram 4.c shows that when the number of components is 7, the model generate the best result.

```{r}
pcr.pred=predict(pcr_model,test.num,ncomp=7) 
cat('MSE for PCR:',mean((pred-test$egalit_scale)^2))
```

## 4.d Partial least squares regression

```{r}
library(pls)
pls_model <- plsr(egalit_scale~., data = train.num, scale = TRUE, validation = "CV")
validationplot(pls_model,val.type = 'MSEP',
               main = 'PLSR Model and its MSE (4.d)',
               ylab = 'MSE by cross-validation')
summary(pls_model)
```
Diagram 4.d shows that when the number of components is 2, the model generate the best result.

```{r}
pls.pred=predict(pls_model,test.num,ncomp=2) 
cat('MSE for PLS:',mean((pred-test$egalit_scale)^2))
```


In question 4, MSE values for all models are in the range of 84-86, which are not significantly different from each other. Thus, I do think these four models' performances are approximately equal. 

# Question 5

## 5.1  Linear regression
```{r}
library(vip)
library(pdp)
library(devtools)
library(iml)

features = train.num %>% dplyr::select(-egalit_scale)
response = as.numeric(as.vector(train.num$egalit_scale))
predictor.lm = Predictor$new(model = lm.fit,data = features, y=response)
plot(Interaction$new(predictor.lm))
```
## 5.2 Elastic Net
```{r}
vip(net.model)
```
## 5.3 PCR
```{r}
predictor.pcr = Predictor$new(model = pcr_model,data = features, y=response)
plot(Interaction$new(predictor.pcr))
```
## 5.4 PLS
```{r}
predictor.pls = Predictor$new(model = pls_model,data = features, y=response)
plot(Interaction$new(predictor.pls))
```

Based on diagrams above, the most importance features for each models are:
Linear regression: income06,tolerance,child
Elastic net regression: income06, tolerance, child
Principal component regression (for ncomps =7): tvhours,income06,con_govt 
Partial least squares regression (for ncomps = 2): wordsum, tolerance, income06

The above results show that income06, tolerance and child are most significant features of the model, I think I get these results because I failed to perform feature standardization.