---
title: "HW4"
output: pdf_document
---
Question 1 Preparation Codes
```{r}

library(ISLR)
library(boot)
library(margins)
library(glmnet)
library(dplyr)
library(tidyr)
library(caret)

set.seed(317)
getwd()
setwd("/Users/nishengwenxin/Desktop")
train = read.csv("gss_train.csv")%>%
  na.omit

test = read.csv("gss_test.csv")%>%
  na.omit

preprocessParams <- preProcess(select_if(train, is.numeric), 
                               method=c("center", "scale"))
train.num <- predict(preprocessParams, select_if(train, is.numeric))

preprocessParams <- preProcess(select_if(test, is.numeric), 
                               method=c("center", "scale"))
test.num <- predict(preprocessParams, select_if(test, is.numeric))
```
# Question 1

Find the optimal degree d.

```{r}
seed = 317
cv.MSE<- rep(NA, 10)
for (i in 1:10) {
    glm.fit <- glm(egalit_scale ~ poly(income06, i), data = train)
    cv.MSE[i] <- cv.glm(train, glm.fit, K=10)$delta[1]}

plot( x = 1:10, y = cv.MSE, xlab = "power of income", ylab = "10-fold CV error", 
      main = 'Choosing the optimal degree d (1.a)',
      type = "b", pch = 19, lwd = 2, bty = "n", 
      ylim = c( min(cv.MSE) - sd(cv.MSE), max(cv.MSE) + sd(cv.MSE) ) )

points( x = which.min(cv.MSE), y = min(cv.MSE),
        col = "red", pch = "X", cex = 1.5 )

```

From diagram 1.a, we can see that the the relationship between the responsible variable $Y_{egalit_scale}$ and the predictive variable $X_{income06}$ in the gss-training data set is best modeled by the polynomial regression of degree 2. Equivalently, we get the optimal model with the expression: $$Y= \beta_0 + \beta_1 X +\beta_2 X^2 + \epsilon $$

It's worth to notice that $d=2$ is an acceptable value because generally, it is unusual to use d greater than 3 or 4 because for large values of d, the polynomial curve can become overly flexible and can take on some very strange shapes.

The diagram below (diagram 1.b) shows the resulting polynomial fit to the data and the AME (diagram 1.c)
```{r}
range.income06 <- range(train$income06)
income06.grid <- seq(from = range.income06[1], to = range.income06[2])

fit <- lm(egalit_scale ~ stats::poly(income06, 2), data = train)
preds <- predict(fit, newdata = list(income06 = income06.grid))
plot(egalit_scale ~ income06, 
     main = 'The polynomial fit with the optimal degree (1.b)',
     ylab = 'Predicted egalitarian scale ',
     data = train, col = "blue")
lines(income06.grid, preds, col = "red", lwd = 2)

cplot(fit,"income06",what = "effect",
      main= "AME for income06 (1.c)")
```



# Question 2
Choose the optimal number of cuts. 
```{r}
cv.MSE <- rep(NA, 10)
# for each cut perform 10-fold cross-validation
for (i in 2:10) {
  train$income06.cut <- cut(train$income06, i)
  lm.fit <-  glm(egalit_scale ~ income06.cut, data = train)
  cv.MSE[i] <-  cv.glm(train, lm.fit, K = 10)$delta[1]
}
  plot(2:10, cv.MSE[-1], xlab = "Number of cuts", ylab = "10-fold CV Error",
       main = 'Choosing the optimal number of cuts (2.a)',
     type = "b", pch = 19, lwd = 2, bty ="n")

  points( x = which.min(cv.MSE), y = min( cv.MSE, na.rm = TRUE), col = "red", pch = "X", cex = 1.5 )

```
 
 Diagram 2.a shows that the test error is minimized when the number of cuts equals 4. Equivalently, we breaks X (income06) into 4 distinctive bins and pick a different constant for each of these 4 bins. Since there exists dummy varaibles, we in fact get 4+1 = 5 distinctive regions. The mathematical expression for this model is $$Y=\beta_0 + \beta_1 C_1(X) + \beta_2 C_2(X) + ...+\beta_4 C_4(X) + \epsilon$$

The diagram below (diagram 2.b) shows the resulting fit with the 4 number of cuts.

```{r}

fit  <- glm(egalit_scale~ cut(income06, 4), data = train)
preds <- predict(fit, data.frame(income06 = income06.grid))

plot(egalit_scale ~ income06, data = train, col = "blue",
     main = 'The step function with the optimal number of cuts (2.b)',
     ylab = 'Predicted egalitarian scale ')
lines(income06.grid, preds, col = "red", lwd = 2)

```

# Question 3 

Select the optimal number of degrees of freedom


```{r}
library(splines)
set.seed(317)
cv.MSE <- rep(NA, 10)
for (i in 3:10) {
    fit <- glm(egalit_scale ~ ns(income06, df = i), data = train)
    cv.MSE[i] <- cv.glm(train, fit, K = 10)$delta[1]
}

plot(3:10, cv.MSE[-c(1, 2)], xlab = "Degrees of freedom", ylab = "10-fold CV error", 
     main = 'Choosing the optimal degree of freedom (3.a)',type = "l")
d.min <- which.min(cv.MSE)
points(which.min(cv.MSE), cv.MSE[which.min(cv.MSE)], col = "red", cex = 2, pch = 20)
```



Diagram 3.a shows that the test error is minimized when the degree of freedom equals 6. This spline model is a continuous degree-6 polynomial with continuity in
derivatives up to degree 5 at each knot. In addition, since it's a natural spline, the function is linear at the tail, in order to combat the danger of high variance at the outer range of the predictors.

The diagram below (diagram 3.b) shows the resulting fit with the 6 df.

```{r}

fit <- glm(egalit_scale ~ ns(income06, df = 6), data = train)
preds <- predict(fit, data.frame(income06 = income06.grid))

plot(egalit_scale ~ income06, data = train, col = "blue",
     ylab = 'Predicted egalitarian scale ',
     main = 'Natural spline with the optimal degree of freedom (3.b)')
lines(income06.grid, preds, col = "red", lwd = 2)

```

# Question 4

## 4.a Linear Regression
```{r}

library(caret)

lm.fit=lm(egalit_scale~.,data=train.num)

pred<-predict(lm.fit,test.num)
print(lm.fit)
cat('MSE Value:', mean((pred-test.num$egalit_scale)^2))
```
## 4.b Elastic Net
```{r}
train.mat <- model.matrix(egalit_scale ~ ., data = train.num)
test.mat <- model.matrix(egalit_scale ~ ., data = test.num)
for (i in seq(0, 1, .1))
{
  seed = 828
  cv.out <- cv.glmnet (train.mat,train.num$egalit_scale,alpha = i)
  bestlam <- cv.out$lambda.min
  model <- glmnet(train.mat,train.num$egalit_scale,alpha=1,
                  lambda=bestlam)
  pred <- predict(model,s = bestlam ,newx = test.mat)
  MSE <- mean((pred-test.num$egalit_scale)^2)
  cat('lambda:',bestlam,'alpha:',i,'MSE:',MSE,"\n")
}
```
Therefore，the optimal model occurs where alpha is 0.3

```{r}
cv.out <- cv.glmnet (train.mat,train.num$egalit_scale,alpha = 0.3)
bestlam <- cv.out$lambda.min
net.model <- glmnet(train.mat,train.num$egalit_scale,alpha=1,
                  lambda=bestlam)
pred <- predict(net.model,s = bestlam ,newx = test.mat)
MSE <- mean((pred-test.num$egalit_scale)^2)
net.model
cat('MSE for Elastic Net:',MSE)

```

## 4.c Principal component regression
```{r}
library(pls)
pcr_model <- pcr(egalit_scale~., data = train.num, scale = TRUE, validation = "CV")
validationplot(pcr_model,val.type = 'MSEP',
               main = 'PCR Model and its MSE (4.c)',
               ylab = 'MSE by cross-validation')
summary(pcr_model)
```
Diagram 4.c shows that when the number of components is 11, the model generate the best result. However, since too many components may overfit the data, it's safe to take number of components as 7.

```{r}
pcr.pred=predict(pcr_model,test.num,ncomp=7) 
cat('MSE for PCR:',mean((pcr.pred-test.num$egalit_scale)^2))
```

## 4.d Partial least squares regression

```{r}
library(pls)
pls_model <- plsr(egalit_scale~., data = train.num, scale = TRUE, validation = "CV")
validationplot(pls_model,val.type = 'MSEP',
               main = 'PLSR Model and its MSE (4.d)',
               ylab = 'MSE by cross-validation')
summary(pls_model)
```
Diagram 4.d shows that when the number of components is 2, the model generate the best result.

```{r}
pls.pred=predict(pls_model,test.num,ncomp=2) 
cat('MSE for PLS:',mean((pls.pred-test.num$egalit_scale)^2))
```


In question 4, MSE values for elastic net, PCR and PLS  are in the range of 0.92-0.93 , which are not significantly different from each other. Thus, I do think these four models' performances are approximately equal. The MSE value for linear regression is slihghtly higher around 0.935, which means the other three models perform slightly better than it does. 

# Question 5

## 5.1.a Linear regression
```{r}
library(vip)
library(pdp)
library(devtools)
library(iml)

features = train.num %>% dplyr::select(-egalit_scale)
response = as.numeric(as.vector(train.num$egalit_scale))

predictor.lm = Predictor$new(model = lm.fit,data = features, y=response)
plot(Interaction$new(predictor.lm))
```
## 5.1.b Elastic Net
```{r}
net = train(
    egalit_scale ~ ., data = train.num, method = "glmnet",
    tuneLength = 10)
net.best = train(
    egalit_scale ~ ., data = test.num, method = "glmnet",
     tuneGrid = expand.grid(alpha = net$bestTune$alpha,
                            lambda = net$bestTune$lambda))

predictor.net = Predictor$new(model = net.best,data = features, y=response)
plot(Interaction$new(predictor.net))

```
## 5.1.c PCR
```{r}
folds = trainControl(method = "cv", number = 10)
pcr = train(
    egalit_scale ~ ., data = train.num, method = "pcr",
    trControl = folds,
    tuneLength = 10)
pcr.best = train(
    egalit_scale ~ ., data = test.num, method = "pcr",
    tuneGrid = expand.grid(ncomp = pcr$bestTune$ncomp))

predictor.pcr = Predictor$new(model = pcr.best,data = features, y=response)
plot(Interaction$new(predictor.pcr))
```
## 5.1.d PLS
```{r}
pls = train(
    egalit_scale ~ ., data = train.num, method = "pls",
    trControl = folds,
    tuneLength = 10)
pls.best = train(
    egalit_scale ~ ., data = test.num, method = "pls",
    tuneGrid = expand.grid(ncomp = pls$bestTune$ncomp))

predictor.pls = Predictor$new(model = pls.best,data = features, y=response)
plot(Interaction$new(predictor.pls))
```

Based on diagrams above, features with the greatest overal interaction strength for each models are:
Linear regression: age,child,sibs
Elastic net regression: sibs, con_govt,income06
Principal component regression : income06, childs, science_quiz
Partial least squares regression: age, income06, wordsum

To sum up, from above four models, the following features have significant overal interaction strength:
INCOME06(3), CHILD (2), SIBS(2), AGE(2)
which means,he variation of the prediction depends heavilyon the interaction of thease features.

# 5.2 Feature Importance

Here's another measurement.

## 5.2.a lm
```{r}
vip(lm.fit)
```
##5.2.b Elastic Net
```{r}
vip(net.best)
```
## 5.2.c PCR
```{r}
vip(pcr.best)
```
## 5.2.d PLS
```{r}
vip(pls.best)
```
From above diagrams, the most important features for each models are:
Linear regression: income06,age,tvhours
Elastic net regression: income06,con_govt,tvhours
Principal component regression (for ncomps =7):income06,con_govt,sibs
Partial least squares regression (for ncomps = 2): income06,con_govt,tvhours

To sum up, from above four models, the following features have significant overal interaction strength:
INCOME06(4),TVHOURS (3), CON_GOVT(3)

The feature importance indicates the variance of the partial dependence functionconditional on the other feature. Higher the variance, stronger the features interaction with each other. THerefore, above three predictors have strong interaction with each other. 